\section{Introduction}\label{intro}
% Occurence of sensors and their data
At the time this paper is written sensors can be found everywhere and their usage is increasing in many fields. Gyroscopes, cameras, compasses, accelerometers and many more are integrated in smartphones. Physical machines are tracking their movement through vibration sensors, health care systems in hospitals visualize the heart beat of a patience and voltmeters measure the generated power in a solar plant. In these examples sensor values are collected over time and time series data (TSD) is produced.\\
% Collections of sensors produce multi variate TSD
In some scenarios the measurements of different sensors are combined. Physical machines sometimes track vibration and motor rotations, health care systems visualize heart beat and blood sugar level and solar plants measure voltage and current.
Collections of different sensors measuring at a common time window produce Multi Variate Time Series Data (MV-TSD).\\
% dependance on MV-TSD
Applications that produce MV-TSD evaluate the data and following decisions or actions depend on a correct analysis.
Normally values change constantly in repetetive patterns. This is when the machine, the patience health or the solar plant is functioning like it is supposed to. But sometimes the values change unpredicted because of differing surroundings or other influences. This can lead to serious situation, especially in the health scenario. When the patience's heart beat changes its pattern the health can of the patience can be endangered. Machine measurements can detect a fault and a solar plant may detect a decline in the generated power.\\
% Importance of Anomaly Detection
Recognising and reacting to these changes in MVTSD can therefore be very important% TODO\citeA.
. But interruptions occur in different forms. In some cases they never occured like this before. This raises the demand for a tool to detect anomalies in time series data without any further knowledge of the anomaly.\\\\
% Difficulty of Anomaly Detection in multi veriate time series data
% TODO
% Why is the literature research important
This paper conducts a detailed literature research to contribute in the development of a anomaly dection tool. It provides an overview of the latest trends in representation learning and extracts the possible solutions addressing the described probelm. The paper focuses on the following research question:\\\\
% Research question
What are the different types of representation learning possible for Zero Shot Anomaly Detection in time series applications?\\\\
% Structure of this paper
A literature research concerning the topic is conducted and the best choices are implemented on a test data set. We begin in \ref{theory} by defining necessary terms and how they are used. In \ref{review} the found papers and their methodologies are presented and explained. The methodologies are compared and evaluated for usability in the context of Zero-Shot Anomaly Detection in \ref{application}. The implementation of suiting techniques is provided in \ref{implementation}. Finally the results are discussed and concluded in \ref{discussion}
