\section{Application for Zero Shot Anomaly Detection on Multi Variate Time Series Data}\label{application}
In this chapter the found articles are filtered using the exclusion criteria defined in \ref{criteria}. By excluding methods that haven't been tested with multiple input variables we answer RQ2. Methods that are not tested in Zero-Shot Learning Scenarios are also excluded which covers RQ4.
More precisely we want to know by the defined filter process which of the proposed RL types are best suited for Zero Shot Anomaly Detection in multi-variate time series data. In this chapter a selection of appropriate methods for Time Series Data Anomaly Detection out of \ref{review} is extracted. Here the priors described in \ref{theory:evaluation} are used in order to rate the RL types.
% TODO put formulas for a time series and the anomalies. maybe some anomaly score too
 % table
 In order to achieve a successful implementation in the next chapter, the focus is on the opensource availability of the described models. Only models which are available and well documented are chosen for further examination. In \ref{tab_rl_methods} the links to public repositories are listed. They are found by citations in the paper.
 \begin{table}
   \caption{Representation learning methodologies and the classification by exclusion criteria. Methods tested with multiple input variables (MV). Methods tested with anomalies which never occured in the training data (ZSL). Links to repositories containing the models are listed under Implementation.}\label{tab_rl_methods}
   \include{includings/methods.tex}
 \end{table}

 % TODO: write something in every column
 % TODO: tick mark instead of github link
 %TODO tick boxes in table

TODO \shortcite{fung_model_2024}  %TODO
%     Spezifische Anwendung in Zero Shot Anomaly Detection
%  Modelle und Architekturen: Detaillierte Beschreibung der spezifischen Modelle und Architekturen, die für Zero Shot Anomaly Detection verwendet werden.
%  Ergebnisse und Erkenntnisse: Diskussion der Ergebnisse aus den relevanten Studien und deren Bedeutung.
