\begin{longtable}[]{@{}llllll@{}}
\toprule
Method Name & Author & Concepts & MV & ZSL & OSA\tabularnewline
\midrule
\endhead
INRAD & Jeong et al. & MLP & \cmark & \xmark & \cmark\tabularnewline
OmniAnomaly & Su et al. & RNN & \cmark & \xmark & \cmark\tabularnewline
CNN based method & Kravchik et al. & CNN & \cmark & \xmark & \xmark\tabularnewline
CNN based method & He et al. & CNN & \cmark & \cmark & \xmark\tabularnewline
SLMR & Miao et al. & CNN & \cmark & \xmark & \xmark\tabularnewline
Debiased CL & Zhang et al. & CL & \cmark & \cmark & \xmark\tabularnewline
CARLA & Darban et al. & CL & \cmark & \cmark & \cmark\tabularnewline
CL-TAD & Ngu et al. & CL & \cmark & \xmark & \cmark\tabularnewline
COCA & Wang et al. & CL, OCC & \cmark & \xmark & \cmark\tabularnewline
CL based method & Lee et al. & CL, OCC & \cmark & \cmark & \xmark\tabularnewline
CL based method & Chen et al. & CL & \cmark & \xmark & \xmark\tabularnewline
TS2Vec & Yue et al. & CL & \cmark & \cmark & \cmark\tabularnewline
TimeAutoAD & Jiao et al. & CL & \cmark & \xmark & \xmark\tabularnewline
ContrastAD & Li et al. & CL & \cmark & \xmark & \xmark\tabularnewline
Dddetector & Yang et al. & CL & \cmark & \xmark & \cmark\tabularnewline
TRL-CPC & Pranavan et al. & CPC & \cmark & \xmark & \xmark\tabularnewline
TiCTok & Kang et al. & CPC & \cmark & \xmark & \xmark\tabularnewline
MGCLAD & Qin et al. & CL & \cmark & \xmark & \cmark\tabularnewline
TriAD & Sun & CL & \xmark & \xmark & \cmark\tabularnewline
AE based method & Provotar et al. & AE, LSTM & \cmark & \xmark &
\xmark\tabularnewline
UCAE-TENN & Nivarthi et al. & UAE, LSTM & \cmark & \cmark & \xmark\tabularnewline
VRAE based method & Pereira et al. & VRAE & \cmark & \cmark &
\xmark\tabularnewline
AE based method & Ramirez et al. & VRAE & \cmark & \cmark & \xmark\tabularnewline
Bi-VRQRAE & Kieu et al. & VRAE & \cmark & \xmark & \cmark\tabularnewline
FuSAGNet & Han et al. & AE, GNN & \cmark & \xmark & \cmark\tabularnewline
MSTVAE & Pham et al. & VAE, CNN & \cmark & \xmark & \cmark\tabularnewline
deep AOC & Mou et al. & AE, OCC & \cmark & \xmark & \cmark\tabularnewline
AE based method & Zhang et al. & AE & \xmark & \cmark & \xmark\tabularnewline
RANSynCoders & Abdulaal et al. & AE & \cmark & \xmark & \cmark\tabularnewline
TCN-AE & Thill et al. & AE, TCN & \cmark & \cmark & \xmark\tabularnewline
MEGA & Wang et al. & AE, GCN & \cmark & \cmark & \cmark\tabularnewline
DAEMON & Chen et al. & AE & \cmark & \cmark & \xmark\tabularnewline
TSMAE & Gao et al. & AE, LSTM & \cmark & \xmark & \xmark\tabularnewline
GRU-AE & Gong et al. & AE, GRU & \cmark & \xmark & \xmark\tabularnewline
MSCVAE & Yokkampon et al. & AE & \cmark & \cmark &\tabularnewline
AE based method & Wang et al. & AE & & &\tabularnewline
MOMENT & Goswami et al. & AE & & \cmark & \cmark\tabularnewline
TranAD & Tuli et al. & T & & & \cmark\tabularnewline
Transformer based method & Ye et al. & T & & &\tabularnewline
AnomalyTransformer & Xu et al. & T & & & \cmark\tabularnewline
DATN & Wu et al. & T & & &\tabularnewline
LLM based method & Li et al. & T & & &\tabularnewline
TiSAT & Doshi et al. & T & & &\tabularnewline
TCF-Trans & Peng et al. & T & & &\tabularnewline
DCT-GAN & Li et al. & T & & &\tabularnewline
Transformer based method & Shin et al. & T & GAN & &\tabularnewline
SL based method & Beggel et al. & SL & & &\tabularnewline
SL based method & Alshaer et al. & SL & & &\tabularnewline
SL based method & Liang et al. & SL & \cmark & &\tabularnewline
IPS & Li et al. & SL & & &\tabularnewline
& Li et al. & C & & &\tabularnewline
& Niu et al. & LSTM & & &\tabularnewline
MMDF & Duan et al. & MMDF & & &\tabularnewline
\bottomrule
\end{longtable}
